### 2016-02-22 Projects/DelCo_2014_2015/Reports/Hose_Stream_Report/TeX_Files/DelCo_Hose_Stream_Tests.tex
Removed unnecessary plot figs, updated script to generate new hose figs

\begin{tabbing}
\hspace{1.5in} \= \\
% FDS \> Fire Dynamics Simulator \\
% HGL \> Hot Gas Layer \\
% HRR \> Heat Release Rate \\
% HRRPUA \> Heat Release Rate per Unit Area \\
BDP \> Bi-Directional Probe \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

@@

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is a fog nozzle. Most fog nozzles contain an adjustable tip that can be rotated to produce different stream patterns and are therefore often referred to as combination nozzles. 
%\textit{Essentials of Fire Fighting} defines 
%https://docs.google.com/presentation/d/1hZ5DZhYXxCkcIEHNZKumH48b_Q53iM2D7asXkMYlVMM/embed?hl=en&size=l&slide=id.p39

%http://www.tft.com/literature/library/files/bfa.pdf
%Keith Royer, WATER FOR FIRE FIGHTING, p. 7
In addition to the type of stream pattern used, fire suppression tactics using hose streams can differ in the method used to apply the hose stream. For example, Nelson and Royer suggest using a clockwise rotation instead of a counterclockwise application pattern for the following reasons:
\begin{enumerate} 
	\item Clockwise rotation is safer since it drive smoke, gases, and flames away from the nozzle. Counterclockwise rotation does just the opposite.
	\item Clockwise rotation produces steam with an active rolling action. Counterclockwise rotation produces steam with an inactive and lazy action.
	\item A clockwise rotation produces a faster knockdown time.
\end{enumerate}

**** better explanation of hose streams with more complete content and information about existing research to be added ****

\section{Objectives}
\label{sec:objectives}
This report is concerned with examining the impact of hose stream and application pattern selection on air movement inside a residential structure. Numerous experiments containing different flow path configurations were conducted to study the impact of hose stream and water application selection on ventilation in a structure. More specifically, the experiments were conducted to fulfill the following objectives:
\begin{enumerate}
	\item Determine which hose stream patterns cause significant air movement in a structure and the amount of air movement they induce compared to other stream patterns
	\item Determine the impact of different water application patterns on the ventilation in a structure
\end{enumerate}

@@

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Pictures/hose_streams.pdf}
	\caption[Hose stream patterns used in experiments.]{The three types of stream patterns used during water flow experiments. From left to right: straight stream, narrow fog stream, and wide fog stream.}
	\label{fig:hose_streams}
\end{figure}
\FloatBarrier

% =========================
% = EXPERIMENTAL OVERVIEW =
% =========================
\chapter{Experimental Overview}
\label{chap:exp_overview}

To study the impact of different hose stream and application pattern selections on air movement inside a structure, various water flow experiments were conducted in two experimental structures designed to replicate typical residential structures. The experiments focused on studying the differences in amount of air movement caused by specific combinations of three different hose stream patterns and four different water application patterns.

\section{Experimental Setup}
\label{sec:exp_setup}

The series of field experiments described in this report were conducted in two structures of similar design located at the Delaware County Emergency Services Training Center in Sharon Hill, PA. Sensors were installed throughout the residential-sized test structures to measure air velocity during the water flow experiments.

@@

\subsection{Instrumentation}
\label{sec:instrumentation}
A discussion of uncertainties for each type of measurement in this report can be found in Section~\ref{sec:uncertainty}. For every experiment, gas velocity at an interior doorway was measured using differential pressure transducers connected to bi-directional velocity probes (BDPs)~\cite{McCaffrey:Combustion_and_Flame}. An array of 6 BDPs located at the northeast interior doorway of the East Structure is pictured below in Fig.~\ref{fig:BDPs}. A single thermocouple was attached to each BDP to measure the temperature to be used in conjunction with the pressure measurement to derive velocity. Every thermocouple attached to the BDPs were exposed-bead, Chromel-Alumel (type K) with a 1.0~mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2~mm (0.13 in) diameter Inconel shield that was 0.76~m (2.5~ft) in length.

The approximate locations of the BDPs used to measure air velocity are annotated in the plan views of the East Structure and West Structure experimental setups (Fig.~\ref{fig:east_setup} and Figs.~\ref{fig:flow_path_1}-\ref{fig:flow_path_2}, respectively) found in Section~\ref{sec:exp_procedure}. To minimize potential measurement errors due to external environmental conditions (i.e., wind), only velocity data measured by BDPs located at interior doorways are considered in this report. The set of BDPs at the northeast interior doorway of the East Structure, denoted as ``A6", contained 6 probes located at distances of 0.30 m, 0.60 m, 0.90 m, 1.20 m, 1.50 m, and 1.80~m below the soffit of the doorway. The set of BDPs at the interior doorway in the West Structure, denoted as ``A10", contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95~m below the soffit of the corresponding doorway.

\begin{figure}[!ht]
	\includegraphics[width=3.5in]{../Figures/Pictures/BDPs_east}
	\caption[Array of BDPs at interior doorway in East Structure.]{Array of BDPs located at the northeast interior doorway in the East Structure. View is looking at the array from Room C.}
	\label{fig:BDPs}
\end{figure}
\FloatBarrier

@@

BDPs with pressure transducers and single thermocouples were used to measure the gas velocity. Bare-bead Type K thermocouples were co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for BDPs of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments is estimated to be $\pm18$\%.

@@

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Pictures/flow_meter}
	\caption[Flow meter used to measure flow rate during experiments.]{Pressure and flow meter combination used to measure water flow rate during the experiments.}
	\label{fig:flow_meter}
\end{figure}
\FloatBarrier

\section{Experimental Procedure}
\label{sec:exp_procedure}
Every experiment discussed in this report involved discharging water at a rate of 120 GPM from a combination nozzle set to flow water in a straight, narrow fog, or wide fog stream pattern. Specific doors were opened or closed before and/or during the experiments to change the ventilation in the structure.

A total of seven test series were conducted to study the impact of different hose stream and application patterns on gas movement in a structure; three test series used a combination nozzle attached to a monitor to flow water and the other three used a combination nozzle attached directly to a hoseline to flow water (Fig.~\ref{fig:monitor+handline}). Experiments that used a monitor to flow water always applied the stream in a fixed position and thus were focused on changing the location of water application in addition to the stream pattern, while the experiments that used a combination nozzle attached directly to a hoseline to flow water focused on varying the pattern in which water was applied in addition to the stream pattern.

\begin{figure}[!ht]
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/monitor}
	\end{center} 
	\endminipage \hfill
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/handline}
	\end{center}
	\endminipage \hfill
	\caption[Monitor and handline equipped with a combination nozzle.]{Monitor (left) and 1.75 in handline (right) equipped with a combination nozzle and used to flow water during experiments.}
	\label{fig:monitor+handline}
\end{figure}
\FloatBarrier

Three different layouts were used for the six test series. The test series conducted in the East Structure, Tests 33 and 34, were performed using an identical structure layout, presented in Fig.~\ref{fig:east_setup}. Test 33 used a monitor and combination nozzle to flow water and Test 34 used a combination nozzle attached directly to a handline to flow water.

The main difference between the two layouts used for the West Structure experiments is the position of the south side door on the first level. Tests 16 and 18 used the configuration with the door opened (Fig.~\ref{fig:flow_path_1}) and Tests 17 and 19 used the configuration with the door closed (Fig.~\ref{fig:flow_path_2}). Additionally, Tests 16 and 17 used a combination nozzle attached to a monitor to flow water, and Tests 18 and 19 used a combination nozzle attached directly to a handline to flow water.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/East_Hose_Test_Annotated}
	\caption[Plan view of the East Structure setup for Tests 33 and 34.]{Plan view of the East Structure setup for Tests 33 and 34. The view is annotated with the approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of BDP A6 (blue square).}
	\label{fig:east_setup}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_18_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 16 and 18.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 16 and 18. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of BDP A10 (blue square). On the second floor, the stairwell door and north side, west double door were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_19_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 17 and 19.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 17 and 19. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of BDP A10 (blue square). On the second floor, the stairwell door and north side, west double door were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_2}
\end{figure}
\FloatBarrier

\subsection{Monitor Experiments}
\label{sec:monitor_procedure}
Numerous experiments were conducted during three different test series using a monitor equipped with a combination nozzle to flow water in a straight stream, narrow fog stream, and wide fog stream pattern. One test series, Test 33, occurred in the East Structure, and the two remaining series, Tests 16 and 17, occurred in the West Structure. Because monitor nozzles are designed to flow water from a set position, water was always applied in a fixed position during the monitor experiments. Thus, the monitor experiments primarily focused on the impact of different hose stream patterns on air movement in a structure.

Test 33 utilized a monitor equipped with a combination nozzle to flow water from Room C of the East Structure to the area above the doorway on the south wall of Room B (Fig.~\ref{fig:test_33_pic}). A plan view of the structure's layout annotated with the approximate location of the BDP array used to measure gas velocity through an interior doorway (A6) and the general direction of the fully established flow path for Test 33 can be found in Fig.~\ref{fig:east_setup}. The procedure for Test 33 was the simplest of all the test procedures. First, water flowed in a straight stream pattern for 30 seconds, then the stream was changed to a narrow fog for 30 seconds, and finally to a wide fog stream for 30 seconds. An identical process was repeated an additional time.

\begin{figure}[!ht]
	\includegraphics[trim=16cm 6.25cm 9cm 6cm, clip=true, width=4in]{../Figures/Pictures/SS_Room_B_Test_33}
	\\~\\
	\includegraphics[trim=16cm 6.25cm 9cm 6cm, clip=true, width=4in]{../Figures/Pictures/NF_Room_B_Test_33}
	\\~\\
	\includegraphics[trim=16cm 6.25cm 9cm 6cm, clip=true, width=4in]{../Figures/Pictures/WF_Room_B_Test_33}
	\caption[Straight stream, narrow fog stream, and wide fog stream during Test 33.]{Water flowing from the monitor in Room C in a straight stream (top), narrow fog stream (middle), and wide fog stream (bottom) pattern aimed at area above the doorway on the south wall of Room B during Test 33.}
	\label{fig:test_33_pic}
\end{figure}

@@

\subsection{Handline Experiments}
\label{sec:handline_procedure}
Numerous experiments were conducted during three different test series using a 1.75 in handline with a combination nozzle to flow water in a straight stream, narrow fog stream, and wide fog stream pattern. One test series, Test 34, occurred in the East Structure, and the remaining series, Tests 18 and 19, occurred in the West Structure. All three test series primarily focused on the impact of different hose stream and water application pattern selections on air movement in a structure.

Test 34 utilized a 1.75 in handline equipped with a combination nozzle to flow water from Room C of the East Structure into Room B (Fig.~\ref{fig:test_34_pic}. The layout of the structure was identical to the layout described above in Section~\ref{sec:monitor_procedure} and shown in Fig.~\ref{fig:east_setup}. The procedure for Test 34 began by flowing water in a straight stream aimed at the ceiling of Room B for 30 seconds in a fixed position. After 30 seconds, the nozzle was rotated in the clockwise direction for 30 seconds. Then, the nozzle was aimed at the south doorway of Room B and water was applied for 30 seconds in a fixed position followed by 30 seconds in a clockwise pattern. This process was repeated for narrow fog and wide fog streams. Then, the entire procedure was repeated for the three types of streams being applied in the fixed and counter-clockwise pattern.

@@

% Where was the handline aimed? Add flow chart? %
Tests 18 and 19 followed identical procedures that involved using a 1.75 in handline equipped with a combination nozzle to flow water from the north side double doors of the West Structure into the first floor room. For every experiment during the two tests, the water stream was aimed at the south side wall of the first floor. The difference between Tests 18 and 19 is that Test 18 contained the same flow path configuration presented in Fig.~\ref{fig:flow_path_1}, while Test 19 contained the same flow path configuration presented in Fig.~\ref{fig:flow_path_2}. Each test series began by using a straight stream to flow water in a fixed position. After 60 seconds of water flow, the stairwell door was opened. One minute later, the north side, west double door on the second floor was opened, and water continued to flow for 60 seconds. Next, the water flow was stopped, the two doors were closed, and the process was repeated three additional times for the sweeping, clockwise, and counterclockwise application patterns. This entire procedure was repeated using narrow fog and wide fog streams. Fig.~\ref{fig:test_18_pic} contains an image of a straight stream pattern being applied during Test 18 with the flow path fully established.

@@

Fig.~\ref{fig:Test_16_A10_Avgs} contains a plot of the average velocity measured by the BDPs at A10 for the straight, narrow fog, and wide fog stream patterns aimed at the near and far targets during Test 16. Similar plots that were generated using measurements obtained from the BDPs at A10 during Test 17 and measurements obtained from the BDPs at A6 during Test 33 are presented in Appendix~\ref{sec:monitor_figs}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Plots/Test_16_West_063014_BDP_A10_stream_avgs}
	\caption{Average velocity measured by the BDPs at A10 during Test 16 for three hose stream patterns.}
	\label{fig:Test_16_BDP_A10_Avg_All}
\end{figure}
\FloatBarrier

For every hose stream and target location combination during the monitor experiments, there was always significant air movement (gas velocity greater than 0.5 m/s) through the interior doorway while the flow path was fully established. Table~\ref{table:all_mon_vel_avgs} lists the average velocity measured by the BDP array at the interior doorway when the flow path was fully established for each hose stream and target location combination studied during Tests 16, 17, and 33.

@@

Fig.~\ref{fig:Test_18_BDP_A10_Avg_All} contains a plot of the average velocity measured by the BDPs at A10 for the straight, narrow fog, and wide fog stream patterns being applied in a fixed position, a sweeping pattern, a clockwise rotation, and counter clockwise rotation during Test 18. Similar plots that were generated using measurements obtained from the BDPs at A10 during Test 19 and measurements obtained from the BDPs at A6 during Test 34 are presented in Appendix~\ref{sec:handline_figs}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Plots/Test_18_West_063014_BDP_A10_stream_avgs}
	\caption{Average velocity measured by the BDPs at A10 during Test 18 for the three hose stream patterns.}
	\label{fig:Test_18_BDP_A10_Avg_All}
\end{figure}
\FloatBarrier

>>>>

\begin{tabbing}
\hspace{1.5in} \= \\
ESTC \> Delaware County Emergency Services Training Center \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

@@

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is a fog nozzle. Most fog nozzles contain an adjustable tip that can be rotated to produce different stream patterns and are therefore often referred to as combination nozzles. The types of streams produced by a fog nozzle are often divided into three categories based on the angle of the stream with respect to the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}, a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}, and a wide fog stream produces an angle between 45\SI{}{\degree} to 80\SI{}{\degree}~\cite{Essentials_FF}. Increasing the angle of the stream allows for a larger area to be covered by water, thus exposing a larger water droplet surface area for heat absorption. However, increasing the angle decreases the distance reached by the stream and its penetration power and increases the potential to disrupt thermal layering increases. 
%\textit{Essentials of Fire Fighting} defines 
%https://docs.google.com/presentation/d/1hZ5DZhYXxCkcIEHNZKumH48b_Q53iM2D7asXkMYlVMM/embed?hl=en&size=l&slide=id.p39

%http://www.tft.com/literature/library/files/bfa.pdf
%Keith Royer, "Water for Fire Fighting", (Engineering Extension Service, Iowa State University, Ames, Iowa) p. 6 
In addition to the type of stream pattern used, fire suppression tactics using hose streams can differ in the method used to apply the hose stream. For example, Royer suggests using a clockwise rotation instead of a counterclockwise application pattern for the following reasons~\cite{Royer}:
\begin{enumerate} 
	\item Clockwise rotation drives most of the heated gases, smoke, and flames away from the nozzle. Counterclockwise rotation does just the opposite.
	\item Steam formed by the clockwise rotation has a violent rolling action. Counterclockwise rotation produces steam with an inactive and lazy action.
	\item A clockweise roation increases the efficiency of water and produces a faster knockdown time than counterclockwise rotation.
\end{enumerate}

\section{Objectives}
\label{sec:objectives}
This report is concerned with examining the impact of hose stream and application pattern selection on air movement inside a residential structure. Numerous experiments containing different flow path configurations were conducted to study the impact of hose stream and water application selection on ventilation in a structure. More specifically, the experiments were conducted to fulfill the following objectives:
\begin{enumerate}
	\item Determine which hose stream patterns cause significant air movement in a structure 
	\item Determine the amount of air movement induced by different stream patterns
	\item Determine the impact of different stream application patterns on the ventilation in a structure
\end{enumerate}

@@

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Pictures/hose_streams.pdf}
	\caption[Hose stream patterns used in experiments.]{The three types of stream patterns used during water flow experiments. From left to right: straight stream, narrow fog stream, and wide fog stream.}
	\label{fig:hose_streams}
\end{figure}

Additionally, a set of experiments comparing the impact of a straight stream from a combination nozzle and a solid stream from a smooth bore nozzle on air flows in a structure was conducted.

% =========================
% = EXPERIMENTAL OVERVIEW =
% =========================
\chapter{Experimental Overview}
\label{chap:exp_overview}
To study the impact of different hose stream and application pattern selections on air movement inside a structure, various water flow experiments were conducted in two experimental structures designed to replicate typical residential structures. The experiments focused on studying the differences in amount of air movement caused by specific combinations of four different hose stream patterns and four different water application patterns.

\section{Experimental Setup}
\label{sec:exp_setup}
The series of field experiments described in this report were conducted in two structures of similar design located at the Delaware County Emergency Services Training Center (ESTC) in Sharon Hill, PA. Sensors were installed throughout the residential-sized test structures to measure air velocity during the water flow experiments.

@@

\subsection{Instrumentation}
\label{sec:instrumentation}
A discussion of uncertainties for each type of measurement in this report can be found in Section~\ref{sec:uncertainty}. For every experiment, gas velocity at an interior doorway was measured using differential pressure transducers connected to bi-directional velocity probes~\cite{McCaffrey:Combustion_and_Flame}. An array of 6 bi-directional probes located at the northeast interior doorway of the East Structure is pictured below in Fig.~\ref{fig:BDPs}. A single thermocouple was attached to each bi-directional probes to measure the temperature to be used in conjunction with the pressure measurement to derive velocity. Every thermocouple attached to the bi-directional probes was exposed-bead, Chromel-Alumel (type K) with a 1.0~mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2~mm (0.13 in) diameter Inconel shield that was 0.76~m (2.5~ft) in length.

The approximate locations of the bi-directional probes used to measure air velocity are annotated in the plan views of the East Structure and West Structure experimental setups (Fig.~\ref{fig:east_setup} and Figs.~\ref{fig:flow_path_1}-\ref{fig:flow_path_2}, respectively) found in Section~\ref{sec:exp_procedure}. To minimize potential measurement errors due to external environmental conditions (i.e., wind), only velocity data measured by bi-directional probes located at interior doorways are considered in this report. The set of bi-directional probes at the northeast interior doorway of the East Structure, denoted as ``A6", contained 6 probes located at distances of 0.30 m, 0.60 m, 0.90 m, 1.20 m, 1.50 m, and 1.80~m below the soffit of the doorway. The set of bi-directional probes at the interior doorway in the West Structure, denoted as ``A10", contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95~m below the soffit of the corresponding doorway.

\begin{figure}[!ht]
	\includegraphics[width=3.5in]{../Figures/Pictures/BDPs_east}
	\caption[Array of bi-directional probes at interior doorway in East Structure.]{Array of bi-directional probes located at the northeast interior doorway in the East Structure. View is looking at the array from Room C.}
	\label{fig:BDPs}
\end{figure}
\FloatBarrier

@@

Bi-directional probes with pressure transducers and single thermocouples were used to measure the gas velocity. Bare-bead Type K thermocouples were co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments is estimated to be $\pm18$\%.

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Pictures/flow_meter}
	\caption[Flow meter used to measure flow rate during experiments.]{Pressure and flow meter combination used to measure water flow rate during the experiments.}
	\label{fig:flow_meter}
\end{figure}
\FloatBarrier

\section{Experimental Procedure}
\label{sec:exp_procedure}
Every experiment discussed in this report involved discharging water at a known flow rate from a combination nozzle set to flow water in a straight, narrow fog, or wide fog stream pattern or from a smooth bore nozzle in a solid stream pattern. Specific doors were opened or closed before and/or during the experiments to change the ventilation in the structure.

A total of seven test series were conducted to study the impact of different hose stream and application patterns on gas movement in a structure; three used a combination nozzle attached to a monitor to flow water at 120~GPM, three used a combination nozzle attached directly to a hoseline to flow water at 120~GPM (Fig.~\ref{fig:monitor+handline}), and one used combination and smooth bore nozzles attached to a monitor to flow water at 180~GPM. Experiments that used a monitor to flow water always applied the stream in a fixed position and thus were focused on changing the location of water application in addition to the stream pattern, while the experiments that used a combination nozzle attached directly to a hoseline to flow water focused on varying the pattern in which water was applied in addition to the stream pattern.

\begin{figure}[!ht]
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/monitor}
	\end{center} 
	\endminipage \hfill
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/handline}
	\end{center}
	\endminipage \hfill
	\caption[Monitor and handline equipped with a combination nozzle.]{Monitor (left) and 1.75~in handline (right) equipped with a combination nozzle and used to flow water during experiments.}
	\label{fig:monitor+handline}
\end{figure}
\FloatBarrier

Three different layouts were used for the seven test series. The test series conducted in the East Structure, Tests 33 and 34, were performed using an identical structure layout, presented in Fig.~\ref{fig:east_setup}. Test 33 used a monitor and combination nozzle to flow water and Test 34 used a combination nozzle attached directly to a handline to flow water.

The main difference between the two layouts used for the West Structure experiments is the position of the south side door on the first level. Tests 16 and 18 used the configuration with the door opened (Fig.~\ref{fig:flow_path_1}) and Tests 17, 19, and 70 used the configuration with the door closed (Fig.~\ref{fig:flow_path_2}). Tests 16 and 17 used a combination nozzle attached to a monitor to flow water; Tests 18 and 19 used a combination nozzle attached directly to a handline to flow water; and Test 70 used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water in a straight and solid stream, respectively.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/East_Hose_Test_Annotated}
	\caption[Plan view of the East Structure setup for Tests 33 and 34.]{Plan view of the East Structure setup for Tests 33 and 34. The view is annotated with the approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of A6 (blue square).}
	\label{fig:east_setup}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_18_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 16 and 18.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 16 and 18. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of A10 (blue square). On the second floor, the stairwell door and north side, west double door were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_19_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 17 and 19.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 17 and 19. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of A10 (blue square). On the second floor, the stairwell door and north side, west double door were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_2}
\end{figure}
\FloatBarrier

\subsection{Monitor Experiments}
\label{sec:monitor_procedure}
Numerous experiments were conducted during three different test series using a monitor equipped with a combination nozzle to flow water at 120~GPM in a straight stream, narrow fog stream, and wide fog stream pattern. One test series, Test 33, occurred in the East Structure, and the two remaining series, Tests 16 and 17, occurred in the West Structure. An additional test series, Test 70, used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water at 180~GPM in a straight stream and solid stream, respectively. Because monitor nozzles are designed to flow water from a set position, water was always applied in a fixed position during the monitor experiments. Thus, the monitor experiments primarily focused on the impact of different hose stream patterns on air movement in a structure.

Test 33 utilized a monitor equipped with a combination nozzle to flow water from Room C of the East Structure to the area above the doorway on the south wall of Room B (Fig.~\ref{fig:test_33_pic}). A plan view of the structure's layout annotated with the approximate location of the bi-directional probes array used to measure gas velocity through an interior doorway (A6) and the general direction of the fully established flow path for Test 33 can be found in Fig.~\ref{fig:east_setup}. The procedure for Test 33 was the simplest of all the test procedures. First, water flowed in a straight stream pattern for 30 seconds, then the stream was changed to a narrow fog for 30 seconds, and finally to a wide fog stream for 30 seconds. An identical process was repeated an additional time.

\begin{figure}[!ht]
	\includegraphics[trim=16cm 6.25cm 9cm 6cm, clip=true, width=4in]{../Figures/Pictures/SS_Room_B_Test_33}
	\\~\\
	\includegraphics[trim=16cm 6.25cm 9cm 6cm, clip=true, width=4in]{../Figures/Pictures/NF_Room_B_Test_33}
	\\~\\
	\includegraphics[trim=16cm 6.25cm 9cm 6cm, clip=true, width=4in]{../Figures/Pictures/WF_Room_B_Test_33}
	\caption[Straight stream, narrow fog stream, and wide fog stream during Test 33.]{Water flowing from the monitor in Room C in a straight stream (top), narrow fog stream (middle), and wide fog stream (bottom) pattern aimed at the area above the doorway on the south wall of Room B during Test 33.}
	\label{fig:test_33_pic}
\end{figure}

@@

Test 70 studied the differences in air movement caused by a solid stream from a smooth bore nozzle and straight stream from a combination nozzle attached to a monitor aimed at the same near and far targets used during Tests 16 and 17. The stairwell door was opened for the entire test series. The test series began by flowing water in a straight stream from a combination nozzle aimed at the near target. Then, the west double door on the second floor was opened. After 60 more seconds of flow, the nozzle was closed to stop flow for 30 seconds. The nozzle was opened for 60 seconds and then closed for 30 seconds two additional times. Then, the monitor was aimed at the far target, and the procedure was repeated. This entire process was repeated using the solid stream from the smooth bore nozzle. Fig.~\ref{fig:test_70_pic} contains images of the straight stream from the combination nozzle and solid stream from the smooth bore nozzle during Test 70.

\begin{figure}[!ht]
	\minipage{2.15in}
	\begin{center}
		\includegraphics[width=2in]{../Figures/Pictures/SS_70}
	\end{center} 
	\endminipage
	\minipage{2.15in}
	\begin{center}
		\includegraphics[width=2in]{../Figures/Pictures/SB_70}
	\end{center}
	\endminipage
	\caption[Straight stream from combination nozzle and solid stream from smooth bore nozzle during Test 70.]{Straight stream (left) from combination nozzle and solid stream (right) from smooth bore nozzle during Test 70.}
	\label{fig:test_70_pic}
\end{figure}
\FloatBarrier

\subsection{Handline Experiments}
\label{sec:handline_procedure}
Numerous experiments were conducted during three different test series using a 1.75~in handline with a combination nozzle to flow water in a straight stream, narrow fog stream, and wide fog stream pattern. One test series, Test 34, occurred in the East Structure, and the remaining series, Tests 18 and 19, occurred in the West Structure. All three test series primarily focused on the impact of different hose stream and water application pattern selections on air movement in a structure.

Test 34 utilized a 1.75~in handline equipped with a combination nozzle to flow water from Room C of the East Structure into Room B (Fig.~\ref{fig:test_34_pic}). The layout of the structure was identical to the layout described above in Section~\ref{sec:monitor_procedure} and shown in Fig.~\ref{fig:east_setup}. The procedure for Test 34 began by flowing water in a straight stream aimed at the ceiling of Room B for 30 seconds in a fixed position. After 30 seconds, the nozzle was rotated in the clockwise direction for 30 seconds. Then, the nozzle was aimed at the south doorway of Room B and water was applied for 30 seconds in a fixed position followed by 30 seconds in a clockwise pattern. This process was repeated for narrow fog and wide fog streams. Then, the entire procedure was repeated for the three types of streams being applied in the fixed and counter-clockwise pattern.

@@

% Where was the handline aimed? Add flow chart? %
Tests 18 and 19 followed identical procedures that involved using a 1.75~in handline equipped with a combination nozzle to flow water from the north side double doors of the West Structure into the first floor room. For every experiment during the two tests, the water stream was aimed at the south side wall of the first floor. The difference between Tests 18 and 19 is that Test 18 contained the same flow path configuration presented in Fig.~\ref{fig:flow_path_1}, while Test 19 contained the same flow path configuration presented in Fig.~\ref{fig:flow_path_2}. Each test series began by using a straight stream to flow water in a fixed position. After 60 seconds of water flow, the stairwell door was opened. One minute later, the north side, west double door on the second floor was opened, and water continued to flow for 60 seconds. Next, the water flow was stopped, the two doors were closed, and the process was repeated three additional times for the sweeping, clockwise, and counterclockwise application patterns. This entire procedure was repeated using narrow fog and wide fog streams. Fig.~\ref{fig:test_18_pic} contains an image of a straight stream pattern being applied during Test 18 with the flow path fully established.

@@

Fig.~\ref{fig:Test_16_A10_Avgs} contains a plot of the average velocity measured by the bi-directional probes at A10 for the straight, narrow fog, and wide fog stream patterns aimed at the near and far targets during Test 16. Similar plots that were generated using measurements obtained from the bi-directional probes at A10 during Test 17 and measurements obtained from the bi-directional probes at A6 during Test 33 are presented in Appendix~\ref{sec:monitor_figs}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Plots/Test_16_West_063014_BDP_A10_stream_avgs}
	\caption{Average velocity through stairwell door during Test 16 for three hose stream patterns.}
	\label{fig:Test_16_BDP_A10_Avg_All}
\end{figure}
\FloatBarrier

For every hose stream and target location combination during the monitor experiments, there was always significant air movement (gas velocity greater than 0.5 m/s) through the interior doorway while the flow path was fully established. Table~\ref{table:all_mon_vel_avgs} lists the average velocity measured by the bi-directional probes array at the interior doorway when the flow path was fully established for each hose stream and target location combination studied during Tests 16, 17, and 33.

@@

Fig.~\ref{fig:Test_18_BDP_A10_Avg_All} contains a plot of the average velocity measured by the bi-directional probes at A10 for the straight, narrow fog, and wide fog stream patterns being applied in a fixed position, a sweeping pattern, a clockwise rotation, and counter clockwise rotation during Test 18. Similar plots that were generated using measurements obtained from the bi-directional probes at A10 during Test 19 and measurements obtained from the bi-directional probes at A6 during Test 34 are presented in Appendix~\ref{sec:handline_figs}.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Plots/Test_18_West_063014_BDP_A10_stream_avgs}
	\caption{Average velocity through stairwell door during Test 18 for the three hose stream patterns.}
	\label{fig:Test_18_BDP_A10_Avg_All}
\end{figure}
\FloatBarrier
